# Data Management 


## Differences 

### [{waldo}](https://waldo.r-lib.org/)

```{r}
waldo::compare(c("a", "b", "c"), c("a", "b"))
df1 <- data.frame(x = 1:3, y = 3:1)
df2 <- tibble::tibble(rev(df1))
waldo::compare(df1, df2)
```

### Vector Item Difference   

```{r}
x <- c(1,2,3,4)
y <- c(2,3,4)
setdiff(x, y)
```
<div class="source">
Source: [Stack Overflow, Dec 2012](https://stackoverflow.com/questions/1837968/how-to-tell-what-is-in-one-vector-and-not-another)
</div>


## Quick Q&A 

### replacement for plyr::rbind.fill 
`dplyr::bind_rows()`  
<div class="source">
Source: [Stack Overflow, Jun 2017](https://stackoverflow.com/questions/44464441/r-is-there-a-good-replacement-for-plyrrbind-fill-in-dplyr)
</div> 


### Set colnames in pipe
``df %>% `colnames<-` newnames``  
<div class="source">
Source: [Stack Overflow, Mar 2016](https://stackoverflow.com/questions/35518218/how-to-set-the-row-names-of-a-data-frame-passed-on-with-the-pipe-operator)
</div> 

```{r}
colnames(cars)
cars |> `colnames<-`(c("one", "two")) |> head()
```
### Remove Columns by Var Type  
`df[,-which(sapply(df, class) == "factor")]`  
`df[, sapply(df, class) != "factor"]`  
`df[,sapply(df, is.numeric)]`  
<div class="source">
Source: [Stack Overflow, Mar 2015](https://stackoverflow.com/questions/28548245/how-to-remove-columns-from-a-data-frame-by-data-type); 
</div>



## Links  

- [Stack Overflow, Oct 2019 | using strsplit in a tibble column](https://stackoverflow.com/questions/58253379/r-using-strsplit-in-a-tibble-column)
- [Stack Overflow, Dec 2014 | Assign variable labels of data frame columns](https://stackoverflow.com/questions/27347548/r-assign-variable-labels-of-data-frame-columns)
- [Stack Overflow, Mar 2017 | using strsplit and subset in dplyr and mutate](https://stackoverflow.com/questions/42565539/using-strsplit-and-subset-in-dplyr-and-mutate)  
- [Stack Overflow, May 2013 | add a dplyr group label as a number](https://stackoverflow.com/questions/50099010/r-add-a-dplyr-group-label-as-a-number)
- [GitHub {bigrquery}, Apr 2018 | Can't join across sources · Issue #219 ](https://github.com/r-dbi/bigrquery/issues/219)

