
# Dates and Times

```{r}
#| message: FALSE 
#| echo: FALSE
library(tidyverse) 
library(knitr)
library(kableExtra)
```


## Date Formats 


::: {.size_75em}

```{r}
#| echo: FALSE
#| column: body
tribble(
  ~String, ~Meaning 
  ,  "%a", "Day of the week, abbreviated (Mon-Sun)"
  ,  "%A", "Day of the week, full (Monday-Sunday"
  ,  "%w", "Day of the week, numeric, 0 = Sunday (0-6)"
  ,  "%e", "Day of month (1-31)"
  ,  "%d", "Day of month (01-31)"
  ,  "%m", "Month, numeric (01-12)"
  ,  "%b", "Month, abbreviated (Jan-Dec)"
  ,  "%B", "Month, full (January-December)"
  ,  "%y", "Year, without century (00-99)"
  ,  "%Y", "Year, with century (0000-9999)"
  ,  "%j", "Day of the Year (001-366)"
  ,  "%U", "Week of year, numeric, starting on Sunday (00-52)"
  ,  "%W", "Week of year, numeric, starting on Monday (00-52)"
  ,  "%x", "Locale-specific date"
  ) %>%
 mutate( 
      Code = paste0('format.Date("2020-12-10", "', String , '")')
    , Output = format.Date("2020-12-10", String)
  ) %>%
  mutate(
      String = cell_spec(String, monospace = TRUE)
    , Code   = cell_spec(Code,   monospace = TRUE)
  ) %>%
  #escape = F --> allows conversion of html code
  #ex make cell grey
  #         escape = F --> NA     (in grey font)
  #         escape = T --> <span style=" color: grey !important;" >NA</span>
  kable(escape = F, label = "Date Formats") %>%
  kable_styling(
     # html_font = "arial" #"mono-space" 
     bootstrap_options = c(
          "striped" #add striped color to rows
        , "hover" #highlight row when hover over it 
        , full_width = FALSE #doesn't have to be full width 
      )
  )
```
:::

<div class="source">
Source: [stat.berkeley | Dates and Times in R Stack Overflow](https://www.stat.berkeley.edu/~s133/dates.html) 
</div>

## Time Formats 

::: {.size_75em}

```{r}
#| echo: FALSE
#| column: body
tribble(
  ~String, ~Meaning 
  ,  "%S", "Second (00-59)"
  ,  "%M", "Minute (00-59)"
  ,  "%l",  "Hour, in 12-hour clock (1-12)"
  ,  "%I", "Hour, in 12-hour clock (01-12)"
  ,  "%p", "am/pm"
  ,  "%H", "Hour, in 24-hour clock (00-23)"
  ,  "%X", "Locale-specific time"
  ,  "%c", "Locale-specific date and time"
  ,  "%z", "Offset from GMT"
  ,  "%Z", "Time zone (character)"
  ) %>%
  mutate( 
      Code = paste0('format.Date(as.POSIXct("2020-12-10 15:30:10", tz = "America/Chicago"), "', String , '")')
    , Output = format.Date(as.POSIXct("2020-12-10 15:30:10", tz = "America/Chicago"), String)
  ) %>%
  mutate(
      String = cell_spec(String, monospace = TRUE)
    , Code   = cell_spec(Code,   monospace = TRUE)
  ) %>%
  #escape = F --> allows conversion of html code
  #ex make cell grey
  #         escape = F --> NA     (in grey font)
  #         escape = T --> <span style=" color: grey !important;" >NA</span>
  kable(escape = F, label = "Date Formats") %>%
  kable_styling(
     # html_font = "arial" #"mono-space" 
     bootstrap_options = c(
          "striped" #add striped color to rows
        , "hover" #highlight row when hover over it 
        , full_width = FALSE #doesn't have to be full width 
      )
  )
```
::: 

<div class="source">
Source: [stat.berkeley | Dates and Times in R Stack Overflow](https://www.stat.berkeley.edu/~s133/dates.html) 
</div>


## Time Zone options in R

The above example uses Central time and so I can use `tz = America/Chicago`; other timezone options can be found using the code below: 

```{r}
#check system time zone
Sys.timezone(location = TRUE)
#other time zone options (only show first 20)
OlsonNames()[1:20]
```


## Links 

- [ggplot2 | Section 9: Scales: Date Time Scales](https://ggplot2-book.org/scale-position.html#date-scales)
- [R-Bloggers | Date Formats in R](https://www.r-bloggers.com/2013/08/date-formats-in-r/)