# Fonts  

## ttf fonts  

### Install Packages 


```{r}
#| eval: FALSE 
install.packages("extrafont")
remotes::install_version("Rttf2pt1", version = "1.3.8")
```

As of August 2022, if you just install `{extrafont}` as is, you will have a problem with importing fonts on Windows and there will be an error saying 'No Font Name'.  To solve this, install version 1.3.8 of `{Rttf2pt1}`.  

<div class="source">
Sources: [Stack Overflow, May 2020 | how to resolve the "No Font Name" issue when importing fonts into R with extrafont](https://stackoverflow.com/questions/61204259/how-can-i-resolve-the-no-font-name-issue-when-importing-fonts-into-r-using-ext) and [extrafont GitHub Issue #88 Jul 2021 | wzbillings comment](https://github.com/wch/extrafont/issues/88#issuecomment-890426514)
</div>


### Download All .ttf Fonts in `C:\Windows\Fonts` Folder

```{r}
#| eval: FALSE 
#will import ALL fonts, takes 5-10 min
extrafont::font_import() 
```


### Download Other Fonts by File 

```{r}
#| eval: FALSE 
extrafont::font_import(paths = "C:/PATH/TO/FOLDER/WITH/TTF", prompt = FALSE) 
```

### Use 

Prior to using fonts in R, need to load fonts in each session (this is like using `library()` or `box::use()` for a specific package) 

```{r}
#| eval = FALSE 
#quiet = FALSE to remove listing of of fonts 
extrafont::Loadfonts(devide = "win", quite = TRUE) 
```

## otf fonts 

### Install Packages 

- issues with `{showtext}` package when in use 
- other ggplot2 fonts are affect when actively using this package, making the fonts smaller and making line breaks much larger

```{r}
#| eval: FALSE 
install.packages("showtext")
library(showtext)
showtext_auto() 
font_add("NewFont", regular = "C://PATH//TO//NewFont.otf")
font_families()
```


```{r}
#| eval: FALSE 
#| echo: FALSE 
showtext_auto() 
font_add("Graphik", regular = "C://Users//meichlersmith//Desktop//Fonts//Graphik.otf")
font_families()
```