# qmd/Rmd 


## Loop with child (sub file)

::: {.grid}

::: {.g-col-6}
**Main-file: `.qmd`** 
:::
<!-- END {.g-col-6} first column -->

::: {.g-col-6}
**Sub-file: `plot_for_state.qmd`**
:::
<!-- END {.g-col-6} second column -->

::: 
<!-- END {.grid} -->


::: {.grid}

::: {.g-col-6 style="border: 1px solid #dcdcdc; font-size: 0.7em;"}

\#\# Individual States 

```{r}
#| eval: false
#| results: asis
# results: asis #outputs text as raw md content 

STATE_vec <- c("Alabama", "Alaska", "Arizona", ...)

#create empty list 
res <- vector(mode = "list", length = 3L)

#run loop 
for (i in 1:length(STATE_vec)) { 
  
  #run the qmd with a specific state, save in list
  res[[i]] <- knitr::knit_child(
      "plot_for_state.qmd"      
    , quiet = TRUE
    , envir = environment()
    )
}

#unlist and paste together with a break 
cat(unlist(res), sep = '\n')
```

\#\# Next Section  

:::
<!-- END {.g-col-6} first column -->


::: {.g-col-6 style="border: 1px solid #dcdcdc; font-size: 0.7em;"}

```{r}
#| eval: false
this_STATE <- STATE_vec[i]  
```

\#\#\#  &#96;r this_STATE&#96;

```{r}
#| eval: false 

all_data %>% 
  filter(STATE == this_STATE) %>% 
  plot_function()
```

:::
<!-- END {.g-col-6} second column -->

::: 
<!-- END {.grid} -->

<div class="source">
Sources: <ul>
  <li>[Stack Overvlow, Aug 2020 | R Markdown: loop with child.rmd](https://stackoverflow.com/questions/63011158/r-markdown-loop-with-child-rmd)</li>
  <li>[Harvey Lieberman, Mar 2022 | knitr::knit_child with Quarto](https://www.harveyl888.com/post/2022-05-12-knit_child_quarto/)</li>
  </ul>
</div>



## `quarto_render()` 

- issue rendering qmd and saving in another folder
- my work around is rendering in same place and moving to another folder and new name, then remove the rendered html 
- links 
    * [Stack Overflow, Jun 2022 | Two problems rendering a qmd file with quarto_render from R](https://stackoverflow.com/questions/72346829/two-problems-rendering-a-qmd-file-with-quarto-render-from-r)
    * [RStudio, Aug 2022 | Output directory in quarto-cli not respected?](https://community.rstudio.com/t/output-directory-in-quarto-cli-not-respected/143762)

```{r}
#| eval: false 
#when bugs are fixed 
# quarto::quarto_render(input = "in.qmd", output_file = "products/output.html")

quarto::quarto_render(input = "ind.qmd", output_file = "output.html")
file.copy(from = "output.html", to = "products/output.html", overwrite = TRUE)
rm("output.html")
```




## Output and Pathways  

- [R Markdown Cookbook | 16.6 The working directory for R code chunks](https://bookdown.org/yihui/rmarkdown-cookbook/working-directory.html) 
- [Stack Overflow | Run R markdown (.Rmd) from inside other R script to produce HTML - Stack Overflow](https://stackoverflow.com/questions/65921907/run-r-markdown-rmd-from-inside-other-r-script-to-produce-html)
- [RStudio Community | save rmarkdown output in different folder - R Markdown](https://community.rstudio.com/t/save-rmarkdown-output-in-different-folder/68192/4)

## Style 

- [Stack Overflow | R markdown: how to change style with internal css?](https://stackoverflow.com/questions/34022558/r-markdown-how-to-change-style-with-internal-css)
- [Stack Overflow | rmarkdown: both portrait and landscape layout in a single PDF](https://stackoverflow.com/questions/25849814/rstudio-rmarkdown-both-portrait-and-landscape-layout-in-a-single-pdf)
- [R Markdown Cookbook | 4.4 Unnumbered sections](https://bookdown.org/yihui/rmarkdown-cookbook/unnumbered-sections.html)
- [Stack Overflow | Increase width of entire HTML Rmarkdown output](https://stackoverflow.com/questions/34906002/increase-width-of-entire-html-rmarkdown-output)
- [Steve Sque | Symbols in LaTeX and HTML](https://www.stevesque.com/symbols/)
- [RStudio Community | Copy to clipboard buttons for code blocks - R Markdown](https://community.rstudio.com/t/copy-to-clipboard-buttons-for-code-blocks/45367)


