
# Functions 

## Correlation

### base R correlation: `cor()`  


```{r}
cor(mtcars[,1:4])
```

If data has NA's in any of the values the `cor()` will results in `NA`.  If you want to remove the NA's when calculating correlation do: 

```
cor(..., use = "complete.obs")
```

<span class="source">
Source: [https://stackoverflow.com/questions/3798998/cor-shows-only-na-or-1-for-correlations-why](https://stackoverflow.com/questions/3798998/cor-shows-only-na-or-1-for-correlations-why)
</span>

### Dot Plots for Multiple Variables: `pairs()`  
 
That chart that plots all variables against eachother as a dot plot when looking to see if variables are correlated with eachother 

```{r}
#lots of variables so only look at first 4
testdf <- mtcars[,1:4]
pairs(testdf, main = "title")
```



## if else 

### Statements 


test expression goes in parenthesis `()` and the statement goes in the curly brackets `{}` 

```
if (test) { statment } 
```


R is a bit finicky with where the brackets go; I get errors when I put `else` on a new line by itself - it wants to have the right brackets before it; `} else` 
```
if (test_A) {
statment #1
} if else (test_B) {
} else {
statment #2 
}
```

### Vectorized 


```
ifelse(test, if_true, if_false)
```

Use case when when have more than 2 options (this is a replacement for nested `ifelse()` functions) 

```
case_when(
    x == A ~ "a"
  , x == B ~ "b"
  , x == C ~ "c"
  , TRUE    ~ NA_character # catchall, default is NA, need to specify *type* of NA
)
```


## Links 

- [R package development | 5 Documenting functions](https://combine-australia.github.io/r-pkg-dev/documenting-functions.html)
- [Stack Overflow | Print dataframe name in function output](https://stackoverflow.com/questions/16742951/print-dataframe-name-in-function-output)
- [Stack Overflow | How to write "descriptions for arguments" in R when coding the wrapper functions?](https://stackoverflow.com/questions/54624408/how-to-write-descriptions-for-arguments-in-r-when-coding-the-wrapper-functions)